
##Peer Assesment 2##

##The Natural events that cause the most harm in Health and Economy.##

###Synopsis###
Natural dissasters are really costly in money, injuries and human lifes. We want to analyze which of these are the ones that cause the worst impact.

###Data Processing###
Explanation of how the data was processed for this study.

Loading the data.

```{r}
StormData = read.csv("repdata-data-StormData.csv")
```

Across the United States, which types of events are most harmful with respect to population health?

I'll measure the population health in terms of the injuries and fatalities.I will asume that the larger these averages per event The most harmful.
To try to solve the previous question I'll calculate the average number of injuries and fatalities per event.

```{r}
fatalities_events = aggregate(FATALITIES~EVTYPE,data=StormData,mean)

injuries_events = aggregate(INJURIES~EVTYPE,data=StormData,mean)
```

After merging these sets I will assume an injury and a fatality are equaly bad, hence I'm ordering and displaying the top 10 events that cause the most of both.
These are the top 10 sources of Fatalities and Injuries:

```{r}
all_health = merge(fatalities_events,injuries_events)

all_health_order = all_health[with(all_health, order(-(FATALITIES+INJURIES))),]

injuries_order = all_health[with(all_health, order(-INJURIES)),]

fatalities_order = all_health[with(all_health, order(-FATALITIES)),]

top_health = head(all_health_order,10)
top_health
```

So assuming that both injuries and fatalities generate the same amount of 'harm' we can observe from this graph that the 5 tallest points are:
```{r, echo=FALSE}
head(top_health)
````
```{r}
library(ggplot2)
ggplot(data=top_health, aes(x=EVTYPE, y=(FATALITIES+INJURIES))) + geom_bar(stat="identity")
```

Across the United States, which types of events have the greatest economic consequences?

I'll measure the economic consequences in terms of the damage caused to  crops and properties.
I will asume that the larger these averages per event The most economical consequences they had
.
To try to solve the previous question I'll calculate the average number of damages per event.

```{r}
cropdmg_events = aggregate(CROPDMG~EVTYPE,data=StormData,mean)

propdmg_events = aggregate(PROPDMG~EVTYPE,data=StormData,mean)
```

Both of these sets are in the same unit (USD) because of this they can be compared equally.
I'm ordering and displaying the top 10 events that cause economical impact.
These are the top 10 sources of Economical Impact:

```{r}
all_economical = merge(cropdmg_events,propdmg_events)

all_economical_order = all_economical[with(all_economical, order(-(CROPDMG+PROPDMG))),]

top_economical = head(all_economical_order,10)
top_economical
```

So assuming that both injuries and fatalities generate the same amount of 'harm' we can observe from this graph that the 5 tallest points are:
```{r, echo=FALSE}
head(top_economical)
````
```{r}
library(ggplot2)
ggplot(data=top_economical, aes(x=EVTYPE, y=(CROPDMG+PROPDMG))) + geom_bar(stat="identity")
```

###Results###

The Most harmful events to Health are Heat Waves, Tropical Storm Gordon, Wildfires, Thunderstorms, Tornadoes and High Wind and Seas.
The most deadly being Tornadoes, and the most Injures with Heat Waves.

The most harmful events in the economical sense are Tropical Storm Gordon,Coastal Erosion, Heavy Rain and Flood,  River Floods, and Dust Storms.
With theTropical Storm Gordon being undoubtedly the worst.

If We had to pick a single worst event with economicals and health implications it'd be Tropical Storm Gordon.

